Machine learning topological phases in real space

### 2020-07-14 10_prb_submission/02_reply_referee/00_article_with_comments/article.tex
Writing table with results

In this last section, we show how the theoretical formalism sketched in section \ref{emergent_information_entropy_wave_functions} can be used to instigate yet another round of machine learning, with a flavor of data compression. As we have argued at the end of section \ref{information_entropy_signatures}, the information entropy signatures can be used as a powerful dimensionality reduction tool: by keeping only the most relevant lattice sites in an information entropy signature, the amount of data needed to characterize the topological phase of an SSH system can be largely decreased with little loss of information. Nevertheless, given that the topological phase of a SSH lattice is a global property and not a local one, it is expected that simply extracting local coordinates of real space eigenvectors may not be the most effective way to condense a lattice's topological information. Here we explore the engineering of new global topological features in real space, and how these global features lead to an effective compression of topological information.

The information entropy signatures in wavevector space can be easily computed from their counterparts in real space (figures \ref{feature_importances_ssh1} and \ref{feature_importances_ssh2}) using equation \eqref{fourier_shannon_wave_function} and assuming that the real space information entropy wave functions are real. The discretized version of equation \eqref{fourier_shannon_wave_function},
\begin{equation}
\label{discrete_fourier_transform}
\hat{\psi}_S[n] = \frac{1}{\sqrt{N}}\sum_{m=0}^{N-1}\psi_S[m]e^{-i\frac{2\pi}{N}nm}
\end{equation}
which is simply the Discrete Fourier Transform (DFT) formula for digital signals, will accomplish this. After computing the norm of each component of the resulting discrete information entropy wave function $\hat{\psi}_S[n]$, we recover the information entropy signatures in wavevector space shown in figures \ref{fourier_feature_importances_ssh1} and \ref{fourier_feature_importances_ssh2}.

>>>>

In this section, we explore the information entropy signatures obtained in section \ref{information_entropy_signatures} from an information-theoretic perspective, in particular as related to data compression. As we have argued at the end of section \ref{information_entropy_signatures}, the information entropy signatures can be used as a powerful dimensionality reduction tool: by keeping only the most relevant lattice sites in an information entropy signature, the amount of data needed to characterize the topological phase of a SSH system can be largely decreased with little loss of information. Nevertheless, given that the topological phase of a SSH system is a global property of the whole lattice, it is natural to expect that it should be possible to engineer global features from real space coordinates. Here we show how the symmetries in the real space information entropy signatures in figure \ref{feature_importances} can be exploited to engineer new global topological features, leading to an effective compression of topological information.

%simply extracting local coordinates of real space eigenvectors may not be the most effective way to condense a lattice's topological %information. Here we explore the engineering of new global topological features in real space, and how these global features lead to an %effective compression of topological information.

%The information entropy signatures in wavevector space can be easily computed from their counterparts in real space (figures %\ref{feature_importances_ssh1} and \ref{feature_importances_ssh2}) using equation \eqref{fourier_shannon_wave_function} and assuming that the %real space information entropy wave functions are real.

Given a real space eigenvector $x[m]$, we can compute its coordinates in wavevector space using the Discrete Fourier Transform (DFT),  
\begin{equation}
\label{discrete_fourier_transform}
\hat{x}[n] = \frac{1}{\sqrt{N}}\sum_{m=0}^{N-1}x[m]e^{-i\frac{2\pi}{N}nm}, \quad n=0,\dots,N-1.
\end{equation}
Since the choice of phase of the real space eigenvectors was such that they were all in $\mathbb{R}^{N}$, the eigenvectors in wavevector space computed from equation \eqref{discrete_fourier_transform} will be Hermitian vectors in $\mathbb{C}^N$. The Hermitian symmetry of the wavevector space eigenvectors manifests itself mathematically in closed lattices with periodic boundary conditions as 
\begin{equation}
\label{hermitian_symmetry}
\hat{x}[\overline{k}] = \hat{x}^*[\overline{N-k}], \quad k = 0,\dots, N-1,
\end{equation}
where  we have used the notation $\overline{k} = k \mod N$.

Equation \eqref{hermitian_symmetry} forks into two natural paths to topological feature engineering. In the first path, we exploit the fact that the real part of $\hat{x}[m]$ is even-symmetric around the reciprocal lattice sites $0$ and $\frac{N}{2}$. This leads us to the Discrete Cosine Transform (DCT) topological features,
\begin{equation}
\label{DCT}
\hat{x}_{c}[n] = x[0] + (-1)^n x[M-1] + \sum_{m=1}^{M-2} 2x[m]\cos\bigg(\frac{\pi}{M-1}nm\bigg), \quad n=0,\dots,M-1, 
\end{equation}
where $M = \frac{N}{2}+1$. The second path capitalizes on the fact that de imaginary part of $\hat{x}[m]$ is odd-symetric around the reciprocal lattice sites $0$ and $\frac{N}{2}$, thus yielding the Discrete Sine Transform (DST) topological features,
\begin{equation}
\label{DST}
\hat{x}_{s}[n+1] = \sum_{m=0}^{M-1} 2x[m+1]\sin\bigg(\frac{\pi}{M+1}(n+1)(m+1)\bigg), \quad n=0,\dots,M-1,
\end{equation} 
where $M = \frac{N}{2}-1$.

We ran several numerical experiments to evaluate how the DCT and DST topological features defined in equations \eqref{DCT} and \eqref{DST} perform. The results obtained in each experiment are listed in table \ref{numerical_results}. These numerical experiments differed both on which topological features were used (whether DCT or DST topological features), as well as which features in real space were used to compute the engineered topological features. The results obtained in the numerical experiments of section \ref{numerical_experiments} are also listed in table \ref{numerical_results} for the sake of reference.

\begin{table}
 \begin{tabular}{||c c c c c||} 
 \hline
 SSH system & Real space feat. & Wavevector space feat. & Val. Hamiltonian acc. & Test Hamiltonian acc. \\ % [0.5ex] 
 \hline\hline
 1 & All & None & 1 & 0.991 \\ 
 \hline
 1 & $\mathcal{S}_1$ & None & 0.985 & 0.985\\
 \hline
 1 & $\mathcal{S}'_1$ & None & 0.985 & 0.986\\
 \hline
 1 & DCT & DCT & 0.997 & 0.989 \\
 \hline
 1 & DST & DST & --- & --- \\
 \hline
 1 & $\mathcal{S}_1$ & $\mathcal{E}_1$ & --- & --- \\ 
 \hline
 1 & $\mathcal{S'}_1$ & $\mathcal{E}'_1$ & --- & --- \\ 
 \hline
 1 & $\mathcal{S}_1$ & $\mathcal{O}_1$ & --- & --- \\ 
 \hline
 1 & $\mathcal{S'}_1$ & $\mathcal{O}'_1$ & --- & --- \\ 
 \hline
 2 & All & None & --- & --- \\ 
 \hline
 2 & $\mathcal{S}_2$ & --- & --- & ---\\
 \hline
 2 & $\mathcal{S}'_2$ & --- & --- & ---\\
 \hline
 2 & DCT & DCT & --- & --- \\
 \hline
 2 & DST & DST & --- & --- \\
 \hline
 2 & $\mathcal{S}_2$ & $\mathcal{E}_2$ & --- & --- \\ 
 \hline
 2 & $\mathcal{S'}_2$ & $\mathcal{E}'_2$ & --- & --- \\ 
 \hline
 2 & $\mathcal{S}_2$ & $\mathcal{O}_2$ & --- & --- \\ 
 \hline
 2 & $\mathcal{S'}_2$ & $\mathcal{O}'_2$ & --- & --- \\ 
 \hline

 \end{tabular}
\caption{\label{numerical_results}Your caption.}
\end{table}

Guaranteed to have Hermitian symmetry with respect to the lattice site at $\frac{N}{2}$. Furthermore, given that the lattices were closed with periodic boundary conditions, the   
%After computing the norm of each component of the resulting discrete information entropy wave function $\hat{\psi}_S[n]$, we recover the %information entropy signatures in wavevector space shown in figures \ref{fourier_feature_importances_ssh1} and %%\ref{fourier_feature_importances_ssh2}.
